---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
library(ggplot2)
unzip(zipfile = "activity.zip")
activity <- read.csv("activity.csv")
activity_complete <- activity[complete.cases(activity),]
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
total_per_day <- aggregate(activity_complete$steps, by=list(activity_complete$date), FUN=sum)
colnames(total_per_day) <- c("date", "total_steps")
print(total_per_day)

hist(total_per_day$total_steps, breaks=10, xlab="Total number of steps per day", main="")
print(paste0("Total steps mean: ",round(mean(total_per_day$total_steps))))
print(paste0("Total steps median: ",round(median(total_per_day$total_steps))))
```


## What is the average daily activity pattern?
```{r, echo=TRUE}
average_interval <- aggregate(activity_complete$steps, by=list(activity_complete$interval), FUN=mean)
colnames(average_interval) <- c("interval", "average_steps")

plot(average_interval$interval, average_interval$average_steps, type="l", xlab="interval", ylab="average steps")

max_interval <- average_interval$interval[max(average_interval$average_steps)==average_interval$average_steps]
print(paste0("Interval with the maximum number of steps: ",paste(max_interval, collapse=",")
              ))
```
## Imputing missing values
```{r,echo=TRUE}

# let's take average per interval and use for filling empty values:

for(i in 1:nrow(activity)){
    if(!is.na(activity$steps[i])){next}
    value = average_interval$average_steps[average_interval$interval==activity$interval[i]]
    activity$steps[i] <- value
}
total_per_day_full <- aggregate(activity$steps, by=list(activity$date), FUN=sum)
colnames(total_per_day_full) <- c("date", "total_steps")
print(total_per_day_full)

hist(total_per_day_full$total_steps, breaks=10, xlab="Total number of steps per day", main="")
print(paste0("Total steps mean: ",round(mean(total_per_day_full$total_steps))))
print(paste0("Total steps median: ",round(median(total_per_day_full$total_steps))))


## Are there differences in activity patterns between weekdays and weekends?
weekday <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
activity$day <- NA
for(i in 1:nrow(activity)){
    activity$day[i] <- if(weekdays(as.Date(activity$date[i])) %in% weekday){"weekday"} else{"weekend"}
}

average_interval_full <- aggregate(activity$steps, by=list(activity$interval, activity$day), FUN=mean)
colnames(average_interval_full) <- c("interval", "day", "average_steps")

qplot(x=interval, y=average_steps, data=average_interval_full, geom="line",facets=day~.)


